---
title: "p8105_hw2_waa2119"
author: "William Anderson"
date: "2022-09-30"
output: github_document
---

```{r, message = FALSE}

library(tidyverse)
library(dplyr)
library(readxl)

```


# Homework 2

## Problem 1
First we will read in the data about entrance and exits for subway stations in NYC

We will retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance then convert the entry variable from character to a logical variable


```{r}
subway_data = 
  read_csv("HW2_Data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%

  janitor::clean_names() %>%
  
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, exit_only, vending, entrance_type, ada) %>%
  
  mutate(entry = recode(entry, "YES" = 1, "NO" = 0))

subway_data
  
```


This data contains information about the different subway lines in New York City, the locations of their stations and the different routes they operate on. This data set also has information on whether the entrance types for the stations, whether the stations contain vendors and whether they are compliant with the Americans with Disabilities Act. 

* The number of rows in this dataset are `r nrow(subway_data)` and the number of columns are `r ncol(subway_data)`

* The number of distinct stations are `r nrow(distinct(subway_data, line, station_name))`


```{r}
names_and_ada = distinct(subway_data, line, station_name, ada)

ADA_compliant = sum(names_and_ada$ada, na.rm = TRUE)
```

* The number of ADA compliant stations is `r ADA_compliant`


* The proportion of stations without vending that allow entrance is given below

```{r}
subway_data %>%
  filter(vending != "YES") %>%
  pull(entry) %>%
mean

```


Now we find the number of stations serving the A train and which of these are ADA compliant. 

```{r}

route_number_and_name = distinct(subway_data, station_name, line, ada, across(contains("route")))

A_stations = filter(route_number_and_name, route1 == "A")

A_stations_with_ADA = filter(A_stations, ada == "TRUE")

```

* The number of stations that serve the A train are `r nrow(A_stations)`

* The number of stations that serve the A train and are ADA compliant are `r nrow(A_stations_with_ADA)`



## Problem 2

Now we will read in an excel file for Mr. Trash Wheel

```{r}

mr_trash_wheel = 
  read_excel("HW2_Data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", skip = 1, na = "", range = cell_cols(1:14)) %>%
  janitor::clean_names() %>%
  na.omit() %>%
  mutate(sports_balls = as.integer(round(sports_balls)))

```


Now we import the professor trash wheel dataset

```{r}

prof_trash_wheel = 
  read_excel("HW2_Data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Professor Trash Wheel", skip = 1, na = "") %>%
  janitor::clean_names() %>%
  na.omit() %>%
  mutate(sports_balls = as.integer(round(sports_balls)))

```


Now we combine Mr. Trash Wheel and Professor Trash Wheel datasets

```{r}
mr_trash_wheel$dumpster = as.double(mr_trash_wheel$dumpster)

trash_wheel = 
  full_join(mr_trash_wheel, prof_trash_wheel, by = "dumpster")

trash_wheel
```

This combined data set contains dumpster data for two different waterfront trash wheels in Baltimore, Maryland. They are operated by a company called Healthy Harbor which aims to use these trash wheels to clean and gather trash floating in the harbor in Baltimore. The combined dumpsters data set contains `r count(trash_wheel)` observations with key variables `r colnames(trash_wheel)`. 

* The total weight of trash collected by Professor Trash Wheel from 2017 - 2020 was `r sum(prof_trash_wheel$weight_tons)` tons.

* The total number of sports balls collected by Mr. Trash Wheel in 2020 was `r sum(mr_trash_wheel$sports_balls[381:452])` balls. 

